---
title: "Lab 02 - Plastic waste"
author: "Alison Yao"
date: "Sept 29, 2021"
output: github_document
---

<!-- https://rstudio-education.github.io/datascience-box/course-materials/lab-instructions/lab-02/lab-02-plastic-waste.html -->

## Warm up

### 1

The `Source` pane is where I open up my Rmarkdown file and code in. The `Console` pane is where I test run my code, use command to view data and receive error messages, so that I can make sure what I write in the Rmarkdown file is correct. The `Environment/History/Git` pane is where I check basic information about the data, see the history of what I've run and interact with GitHub. The `File/Plots/Packages/Help/Viewer` pane is where I can see and open up files in the directory, install packages, check R commands and resources, and view the file after knitting. 

### 2

There are 240 observations and 10 variables. 

### 3

NA means that the cell is empty or has no value.

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read_csv("data/plastic-waste.csv")
glimpse(plastic_waste)
```

## Exercises

Before doing the exercises, let's first follow the instructions and work on something else instead.

```{r plastic_waste_per_cap-histogram}
ggplot(plastic_waste, 
       aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)
```

Let's check the one country that is all the way to the right in the histogram.

```{r}
plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```


### Exercise 1

First, we use histograms to plot the distribution of plastic waste per capita faceted by continent. 

```{r plastic-waste-continent-facet}
ggplot(plastic_waste, 
       aes(x = plastic_waste_per_cap, 
           fill = continent)) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(. ~ continent, nrow = 3) + 
  labs(
    title = "Histograms of Plastic Waste per capita Faceted by Continent",
    x = "Plastic Waste per capita", 
    y = "Frequency"
  )
```

By looking at the plot, it seems that all continents apart from North America have plastic waste per capita within 0 and 1. Oceania and South America have lower plastic waste per capita overall than the other 4 continents.

Following the instructions, another way of visualizing numerical data is using density plots.

```{r plastic-waste-continent-density}
ggplot(plastic_waste, 
       aes(x = plastic_waste_per_cap,
           color = continent,
           fill = continent)) + 
  geom_density(alpha = 0.7)
```

I personally think putting all density plots in one is messy, so I prefer to keep them separated. 

```{r plastic-waste-density-facet}
ggplot(plastic_waste, 
       aes(x = plastic_waste_per_cap,
           color = continent,
           fill = continent)) + 
  geom_density(alpha = 0.7) + 
  facet_wrap(~ continent, nrow = 3) + 
  labs(
    title = "Density Plots of Plastic Waste per capita Faceted by Continent",
    x = "Plastic Waste per capita", 
    y = "Density"
  )
```
Now, I think this one looks better.


### Exercise 2

Here, I use a lower `alpha` calue, so that the density plots are more transparent and a bit easier to compare. (Still think the plot above is better, but that is not what the exercise is asking for...)

```{r plastic-waste-density}
ggplot(plastic_waste, 
       aes(x = plastic_waste_per_cap,
           color = continent,
           fill = continent)) + 
  geom_density(alpha = 0.1) + 
  labs(
    title = "Density Plots of Plastic Waste per capita",
    x = "Plastic Waste per capita", 
    y = "Density"
  )
```

### Exercise 3

Because `color` and `fill` are mapped to `continent` variable, while `alpha` is set to a fixed number. The former is mapping, so we use `aes()` function, while the latter is setting, so we use a plotting geom. 

### Exercise 4

First, we visualize using boxplot according to the instructions.

```{r plastic-waste-boxplot}
ggplot(plastic_waste,
       aes(x = continent,
           y = plastic_waste_per_cap)) +
  geom_boxplot()
```

Then, we convert the plot into a violin plot to compare.

```{r plastic-waste-violin}
ggplot(plastic_waste, 
       aes(x = continent,
           y = plastic_waste_per_cap)) + 
  geom_violin()
```

Violin plots give us the distributions, while box plots cannot. Box plots are explicit about the median, 25% and 75% quartiles, upper and lower adjacent values, and outliers, while violin plots are not.  

### Exercise 5

First, we visualize the relationship between plastic waste per capita and mismanaged plastic waste per capita using a scatterplot.

```{r plastic-waste-mismanaged}
ggplot(plastic_waste, 
       aes(x = plastic_waste_per_cap, 
           y = mismanaged_plastic_waste_per_cap)) + 
  geom_point() +
  labs(
    title = "Mismanaged Plastic Waste per capita vs Platic Waste per capita", 
    x  = "Platic Waste per capita",
    y = "Mismanaged Plastic Waste per capita"
  )
```

From the scatterplot, one can see that most points are on or between two straight lines which intersect at the origin, indicating that the two features might be linearly associated. 


### Exercise 6

Then, we color the point according to continents.

```{r plastic-waste-mismanaged-continent}
ggplot(plastic_waste, 
       aes(x = plastic_waste_per_cap, 
           y = mismanaged_plastic_waste_per_cap,
           color = continent)) + 
  geom_point() +
  labs(
    title = "Mismanaged Plastic Waste per capita vs Plastic Waste per capita", 
    x  = "Platic Waste per capita",
    y = "Mismanaged Plastic Waste per capita"
  )
```

Africa and Oceania have a clearer distinction from the others. Most of their points are on the line with the bigger slope, meaning that the mismanagement problem of plastic waste per capita is more severe on these two continents. In contrast, Asia, Europe, North America and South America have more dispersion, meaning that there are some countries on these continents that manage their plastic waste well, but some manage poorly. 

### Exercise 7

First, we make the scatter plots of the relationship between plastic waste per capita and total population and the relationship between plastic waste per capita and coastal population.

```{r plastic-waste-population-total}
ggplot(plastic_waste,
       aes(x = plastic_waste_per_cap, 
           y = total_pop)) + 
  geom_point() + 
  labs(
    title = "Plastic Waste per capita vs Total Population",
    x = "Plastic Waste per capita",
    y = "Total Population"
  )
```

```{r plastic-waste-population-coastal}
ggplot(plastic_waste,
       aes(x = plastic_waste_per_cap, 
           y = coastal_pop)) + 
  geom_point() + 
  labs(
    title = "Plastic Waste per capita vs Coastal Population",
    x = "Plastic Waste per capita",
    y = "Coastal Population"
  )
```

There does not seem to exist a strong linear association in either of the plot. But perhaps it is because we did not zoom in enough due to outliers?

Let's zoom in by filtering the region where the points are too clustered together.

```{r plastic-waste-population-total-zoomin}
plastic_waste %>%
  filter(plastic_waste_per_cap < 1 & total_pop < 1e+08) %>%
  ggplot(aes(x = plastic_waste_per_cap, 
             y = total_pop)) + 
    geom_point() + 
    labs(
      title = "Plastic Waste per capita vs Total Population Zoomed In",
      x = "Plastic Waste per capita",
      y = "Total Population"
    )
```

```{r plastic-waste-population-coastal-zoomin}
plastic_waste %>%
  filter(plastic_waste_per_cap < 1 & coastal_pop < 1e+08) %>%
  ggplot(aes(x = plastic_waste_per_cap, 
             y = coastal_pop)) + 
    geom_point() + 
    labs(
      title = "Plastic Waste per capita vs Coastal Population Zoomed In",
      x = "Plastic Waste per capita",
      y = "Coastal Population"
    )
```

Still,  I don't really see a strong linear association. 

### Exercise 8

In an attempt to recreate the plot, I used a filter to find the right range for the plot. I calculated the ratio of coastal/total population as the x values. I used color blind friendly color to represent different continents. To add a regression line and its confidence interval, I used two `geom` functions and changed the color for `geom_smooth` to black. One thing to note is that I needed to put `color = continent` in `geom_point` instead of `ggplot` like I always do because I need the regression line to fit all datapoints. I added labels and a theme accordingly and removed the border. 

```{r recreate-viz}
plastic_waste %>%
  filter(plastic_waste_per_cap < 1) %>%
  ggplot(aes(x = coastal_pop / total_pop,
             y = plastic_waste_per_cap)) + 
    geom_point(aes(color = continent)) + 
    geom_smooth(color = 'black') +
    labs(
      title = 'Plastic waste vs coastal population proportion',
      subtitle = "by continent",
      x = "Coastal population proportion (Coastal / total population)",
      y = "Plastic waste per capita",
      color = "Continent"
    ) + 
  scale_colour_viridis_d() +
  theme_bw() + 
  theme(
    panel.border = element_blank()
  )
```
