---
title: "EDA"
author: "Alison Yao, Oscar Bray"
date: "11/22/2021"
output: distill::distill_article
---

```{r setup}
library(tidyverse)
library(brms)
library(plotly)
library(corrplot)
```

## Load Dataset

First, read the 3 csv files downloaded from 2019 OECD study of violence against women.

 - Attitudes toward violence: The percentage of women who agree that a husband/partner is justified in beating his wife/partner under certain circumstances 
 
 - Prevalence of violence in the lifetime: The percentage of women who have experienced physical and/or sexual violence from an intimate partner at some time in their life
 
 - Laws on domestic violence: Whether the legal framework offers women legal protection from domestic violence Laws on domestic violence are presented as values ranging from 0 to 1, where 0 means that laws or practices do not discriminate against women’s rights and 1 means laws or practices fully discriminate against women’s rights.

```{r load-data}
attitude_df = read.csv('./Data/attitude.csv')
law_df = read.csv('./Data/laws.csv')
prevalence_df = read.csv('./Data/prevalence.csv')
```

## Initial Findings

By just eyeballing the data, I can see that:
1. Not all countries are listed and not all countries have record on all three features.
2. The range of three features are different (0-100% or 0-1). Strange thing is that law only has 4 values: 0.25, 0.5, 0.75 and 1. It looks more categorical than continuous. 
3. The countries are abbreviated...so I need some way to match the code with the names

```{r load-code}
# Data Source: https://www.iban.com/country-codes
# country_code_df = read.csv('./Data/country_code.csv')
```

```{r}
# check_code <- function(df){
#    for (code in df$LOCATION){
#     if (code %in% country_code_df$Abb3) {
#     } else {
#       print(code, 'not found!')
#     }
#   }
# }
# check_code(attitude_df)
# check_code(law_df)
# check_code(prevalence_df)
# # There should not be any output - Awesome!
```

<!-- So now I have a way to match the country codes to their English names.  -->

Wait a second, I recall that we used similar data in Lab06. And it also provides continent data. This is even better.

```{r}
covid_data <- read_csv("https://wzb-ipi.github.io/corona/df_full.csv")
covid_subset <- covid_data[c('geoid2', 'country', 'continent', 'gdp_pc')]
covid_subset <- covid_subset[!duplicated(covid_subset), ]
covid_subset
# check if there are duplicated countries by running
# dplyr::count(covid_subset, geoid2, sort = TRUE)
```

```{r}
check_code <- function(df){
   for (code in df$LOCATION){
    if (code %in% covid_subset$geoid2) {
    } else {
      print(paste0("Not Found: ", code))
    }
  }
}
check_code(attitude_df)
check_code(law_df)
check_code(prevalence_df)
# There should not be any output - but there are two
```
Let's add them manually.

```{r}
covid_subset <- covid_subset %>% add_row(geoid2 = "HKG", 
                         country = "Hong Kong",
                         continent = "Asia", 
                         gdp_pc = NA)
covid_subset <- covid_subset %>% add_row(geoid2 = "TKM", 
                         country = "Turkmenistan",
                         continent = "Asia", 
                         gdp_pc = NA)
colnames(covid_subset) <- c('Country', 'CountryName', 'Continent', 'GDP_per_capita')
covid_subset
```

Now we can analyze patterns across continents if we want to. 

## Join Data

I will put everything in a dataframe by full outer joining three subsets of the dataframes.

```{r}
colnames(attitude_df) <- c('LOCATION', 'INDICATOR', 'SUBJECT', 'MEASURE', 'FREQUENCY', 'TIME', 'Value', 'Flag.Codes')
attitude_sub <- attitude_df[c('LOCATION', 'Value')]
colnames(attitude_sub) <- c('Country', 'Attitude')

colnames(prevalence_df) <- c('LOCATION', 'INDICATOR', 'SUBJECT', 'MEASURE', 'FREQUENCY', 'TIME', 'Value', 'Flag.Codes')
prevalence_sub <- prevalence_df[c('LOCATION', 'Value')]

colnames(law_df) <- c('LOCATION', 'INDICATOR', 'SUBJECT', 'MEASURE', 'FREQUENCY', 'TIME', 'Value', 'Flag.Codes')
law_sub <- law_df[c('LOCATION', 'Value')]
colnames(law_sub) <- c('Country', 'Law')

colnames(prevalence_sub) <- c('Country', 'Prevalence')
df <- full_join(attitude_sub, prevalence_sub, by = "Country")
df <- full_join(df, law_sub, by = "Country")
df <- left_join(df, covid_subset, by = 'Country')
df
```

Looks good! But there must be many empty values. Let's check.

```{r}
summary(df)
```

Indeed, there are 11 empty values in Attitude and 34 in Prevalence. 

## Plots

Although we already have some amazing interactive visualization on the data source website, we need to do a bit more vis here.

### Single Variable

```{r}
df %>% 
  ggplot(aes(x = Attitude)) +
  geom_histogram(binwidth=2) + 
  labs(
    title = 'Histogram of Attitudes toward violence',
    subtitle = '152 countries included',
    x = 'Attitudes',
    y = 'Frequency'
  )
```

Positively skewed. 

```{r}
df %>% 
  ggplot(aes(x = Attitude)) +
  facet_wrap(. ~ Continent, scales = "free") +
  geom_histogram(binwidth=2) + 
  labs(
    title = 'Histogram of Attitudes toward violence',
    subtitle = 'faceted by 152 countries',
    x = 'Attitude',
    y = 'Frequency'
  )
```

```{r}
df %>% 
  ggplot(aes(x = Prevalence)) +
  geom_histogram(binwidth=2) + 
  labs(
    title = 'Histogram of Prevalence of violence in the lifetime',
    subtitle = '129 countries included',
    x = 'Prevalence',
    y = 'Frequency'
  )
```

Positively skewed. 

```{r}
df %>% 
  ggplot(aes(x = Prevalence)) +
  facet_wrap(. ~ Continent) +
  geom_histogram(binwidth=2) + 
  labs(
    title = 'Histogram of Prevalence of violence in the lifetime',
    subtitle = 'faceted by 129 countries',
    x = 'Prevalence',
    y = 'Frequency'
  )
```

```{r}
df %>% 
  ggplot(aes(x = Law)) +
  geom_bar() + 
  scale_x_continuous(breaks = c(0.25, 0.5, 0.75, 1.00)) + 
  geom_text(
      aes(label=after_stat(count)),
      stat = 'count',
      nudge_y = 3,
      va = 'bottom'
  ) +
  labs(
    title = 'Bar Chart of Laws on domestic violence',
    subtitle = '163 countries included',
    x = 'Law',
    y = 'Frequency'
  )
```

This one is not so informative because there are only 4 values. Are we really treating it as continuous?

```{r}
df %>% 
  ggplot(aes(x = Law)) +
  facet_wrap(. ~ Continent) +
  geom_bar() + 
  scale_x_continuous(breaks = c(0.25, 0.5, 0.75, 1.00)) + 
  geom_text(
      aes(label=after_stat(count)),
      stat = 'count',
      nudge_y = 3,
      va = 'bottom'
  ) +
  labs(
    title = 'Bar Chart of Laws on domestic violence',
    subtitle = 'faceted by 163 countries',
    x = 'Law',
    y = 'Frequency'
  )
```

### Multi-variable

```{r}
df %>% 
  ggplot(aes(x = Attitude, 
             y = Prevalence, 
             color = as.factor(Law),
             shape = Continent)) +
  geom_point() + 
  labs(
    title = 'Scatterplot of Prevalence vs Attitude',
    x = 'Attitude',
    y = 'Prevalence',
    color = 'Law'
  ) + 
  scale_colour_viridis_d()
```

```{r}
df %>% 
  ggplot(aes(x = Attitude, 
             y = Prevalence, 
             color = Continent,
             size = Law)) +
  geom_point(alpha = 0.8) + 
  labs(
    title = 'Scatterplot of Prevalence vs Attitude',
    x = 'Attitude',
    y = 'Prevalence',
    color = 'Continent'
  ) + 
  scale_colour_viridis_d()
```


```{r}
cormat <- round(cor(df[c('Attitude', 'Law', 'Prevalence')] %>% drop_na()), 2)
corrplot(cormat, 
         # method = 'color',
         # order = 'alphabet'
  )
```

```{r}
df %>% 
  ggplot(aes(x = Attitude, y = Prevalence)) +
  geom_point() + 
  # stat_smooth(method="lm") +
  labs(
    title = 'Scatterplot of Prevalence vs Attitude',
    x = 'Attitude',
    y = 'Prevalence'
  )
```

We can see a positive correlation. 

```{r}
# box plots are not as good as bar charts with error bars :(
df %>% 
  ggplot(aes(x = as.factor(Law), y = Prevalence)) +
  geom_boxplot() + 
  labs(
    title = 'Boxplot of Prevalence vs Law',
    x = 'Law',
    y = 'Prevalence'
  )
```

```{r}
# summary statistics are not bad
df %>% 
  ggplot(aes(x = as.factor(Law), y = Prevalence)) +
   stat_summary() + 
  labs(
    title = 'Summary Statistics of Prevalence vs Law',
    x = 'Law',
    y = 'Prevalence'
  )
```


```{r}
df %>% 
  ggplot(aes(x = Law, y = Prevalence)) +
  geom_point() + 
  # stat_smooth(method="lm") +
  labs(
    title = 'Scatterplot of Prevalence vs Law',
    x = 'Law',
    y = 'Prevalence'
  )
```

```{r}
df %>% 
  ggplot(aes(x = as.factor(Law), y = Attitude)) +
  geom_boxplot() + 
  labs(
    title = 'Boxplot of Attitude vs Law',
    x = 'Law',
    y = 'Attitude'
  )
```

```{r}
df %>% 
  ggplot(aes(x = as.factor(Law), y = Attitude)) +
   stat_summary() + 
  labs(
    title = 'Summary Statistics of Attitude vs Law',
    x = 'Law',
    y = 'Attitude'
  )
```

```{r}
df %>% 
  ggplot(aes(x = Law, y = Attitude)) +
  geom_point() + 
  # stat_smooth(method="lm") +
  labs(
    title = 'Scatterplot of Attitude vs Law',
    x = 'Law',
    y = 'Attitude'
  )
```

Now, this is definitely categorical. 

```{r}
plot_ly(x=df$Attitude,
        y=df$Prevalence,
        z=df$Law,
        type="scatter3d")
```

This looks fancy but is pretty useless. I cannot see much from this 3D plot. I can't see any obvious regression plane. The model is probably not going to have a high R-squared. 

## Regression (Not really useful now that I think about it)

Since we cannot see much from the plots, let's do a regression analysis to quantify the relationship.  

I think the y should be prevalence, because it seems to be the result of attitude and law. Let's put the prevalence on the left hand side of the equation.

Now I think they are just three indicators. There is no need to run a regression across them. 

```{r}
# reg1 <- brm(formula = Prevalence ~ Attitude + Law,
#                        data=df, 
#                        refresh = 0,
#                        seed = 123) # stabilize the outcome for reproducibility
# summary(reg1)
```


## Alcohol (Total) - Spurious Association

```{r}
alcohol = read.csv('./Data/alcohol.csv')
```

```{r}
colnames(alcohol) <- c('CountryName', 'Country', 'Alcohol_pc')
df <- left_join(df, alcohol, by = 'Country')
```

```{r}
df %>% 
  ggplot(aes(x = Alcohol_pc, y = Prevalence)) +
  geom_point() + 
  stat_smooth(method="lm") + 
  labs(
    title = 'Prevalence vs Alcohol Consumption per capita',
    x = 'Alcohol Consumption per capita',
    y = 'Prevalence'
  )
```


```{r}
reg2 <- brm(formula = Prevalence ~ Alcohol_pc,
                       data=df, 
                       refresh = 0,
                       seed = 123) # stabilize the outcome for reproducibility
summary(reg2)
```

## poverty_headcount_ratio - Potential Confounder

```{r}
poverty_headcount_ratio  = read.csv('./Data/poverty_cleaned.csv')
```

```{r}
df <- left_join(df, poverty_headcount_ratio, by = 'Country')
```

```{r}
reg3 <- brm(formula = Prevalence ~ Alcohol_pc + poverty_headcount_ratio,
                       data=df, 
                       refresh = 0,
                       seed = 123) # stabilize the outcome for reproducibility
summary(reg3)
```

The absolute value, or the magnitude, of the coefficient of alcohol_pc has dropped from 1.52 to 0.96 (about 37%), so we can conclude that poverty_headcount_ratio is a confounder.

And there are not too many nan's for the poverty_headcount_ratio! Awesome! We found one!

```{r}
summary(df[c('Prevalence','poverty_headcount_ratio')])
```

## Causal Diagram

```{r}
library(dagitty)
library(ggdag)
```

```{r}
vaw <- dagitty("dag{
        VaW -> Law;
        VaW -> Attitude;
        VaW -> Prevalence;
        AlcoholConsumption -> Prevalence;
        PovertyHeadcount -> Prevalence;
        PovertyHeadcount -> AlcoholConsumption;
        Alcohol -> AlcoholConsumption;
        Poverty -> PovertyHeadcount;
}")

ggdag(vaw)
```

The text is too long... How can I fix it???

```{r}
vaw_dag <- dagify(Law ~ vaw,
       Attitude ~ vaw,
       Prevalence ~ vaw + ph + ac + Law + Attitude,
       ac ~ a + ph,
       ph ~ p,
       labels = c("vaw" = "VaW", 
                  "Law" = "Law",
                  "Attitude" = "Attitude",
                  "Prevalence" = "Prevalence",
                  "ph" = "Poverty\n Headcount",
                  "p" = "Poverty",
                  "ac" = "Alcohol\n Consumption",
                  "a" = "Alcohol"
                  ),
       latent = "vaw")

ggdag(vaw_dag, text = FALSE, use_labels = "label")
```

Q: What are adjustment sets???

```{r}
# find variables that aren't related
impliedConditionalIndependencies(vaw_dag)

# find the adjustment set
adjustmentSets(vaw_dag, exposure="ac", outcome = "Prevalence")

# ggdag version
ggdag_adjustment_set(vaw_dag, exposure="ac", outcome = "Prevalence")

```

```{r}
df %>% 
  ggplot(aes(x = poverty_headcount_ratio, y = Prevalence)) +
  geom_point() + 
  stat_smooth(method="lm") + 
  labs(
    title = 'Prevalence vs poverty_headcount_ratio',
    x = 'poverty_headcount_ratio',
    y = 'Prevalence'
  )
```


## - Alcohol use by gender!
```{r load-data}
MaleAlcoholUse <- read_csv('./Data/MaleAlcoholUse.csv')
head(MaleAlcoholUse)
colnames(MaleAlcoholUse) <- c('X1', 'Country', 'AlcoholM_pc')
MaleAlcoholUse <- MaleAlcoholUse[c('Country', 'AlcoholM_pc')]
```
```{r load-data}
FemaleAlcoholUse <- read_csv('./Data/FemaleAlcoholUse.csv')
head(FemaleAlcoholUse)
colnames(FemaleAlcoholUse) <- c('X1', 'Country', 'AlcoholF_pc')
FemaleAlcoholUse <- FemaleAlcoholUse[c('Country', 'AlcoholF_pc')]
```
```{r join-data}
df <- left_join(df, FemaleAlcoholUse, by = 'Country')
df <- left_join(df, MaleAlcoholUse, by = 'Country')
head(df)
```

NOw it's time to graph things!
```{r scatterplot-prevalence-male-alcohol}
df %>% 
  ggplot(aes(x = AlcoholM_pc, y = Prevalence)) +
  geom_point() + 
  stat_smooth(method="lm") + 
  labs(
    title = 'Prevalence vs Male Alcohol Consumption per capita',
    x = 'Alcohol Consumption by men per capita',
    y = 'Prevalence'
  )
```
```{r scatterplot-prevalence-female-alcohol}
df %>% 
  ggplot(aes(x = AlcoholF_pc, y = Prevalence)) +
  geom_point() + 
  stat_smooth(method="lm") + 
  labs(
    title = 'Prevalence vs Female Alcohol Consumption per capita',
    x = 'Alcohol Consumption by women per capita',
    y = 'Prevalence'
  )
```
```{r reg-uni-FemaleAlcohol}
reg_uni2 <- brm(formula = Prevalence ~ AlcoholF_pc,
                       data=df, 
                       refresh = 0,
                       seed = 123) # stabilize the outcome for reproducibility
summary(reg_uni2)
```
```{r reg-uni-MaleAlcohol}
reg_uni3 <- brm(formula = Prevalence ~ AlcoholM_pc,
                       data=df, 
                       refresh = 0,
                       seed = 123) # stabilize the outcome for reproducibility
summary(reg_uni3)
```


```{r reg-multi-FemaleAlcohol}
reg3 <- brm(formula = Prevalence ~ AlcoholF_pc + poverty_headcount_ratio,
                       data=df, 
                       refresh = 0,
                       seed = 123) # stabilize the outcome for reproducibility
summary(reg3)
```

```{r reg-multi-MaleAlcohol}
reg3 <- brm(formula = Prevalence ~ AlcoholM_pc + poverty_headcount_ratio,
                       data=df, 
                       refresh = 0,
                       seed = 123) # stabilize the outcome for reproducibility
summary(reg3)
```
