---
title: "Lab 01 - Hello R"
author: "Alison Yao"
date: "Sept 24, 2021"
output: 
  html_document: 
    theme: journal
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
```

## Exercises

### Exercise 1

The `datasaurus_dozen` file has 1846 rows and 3 columns. There are 3 variable: `dataset`, `x`, and `y`. `dataset` indicates which dataset the data is from; `x` means the x values; `y` means the y values.  

Following the instructions, we can check the frequency table:

```{r dataset-frequency-table}
datasaurus_dozen %>%
  count(dataset) %>%
  print(13)
# strange that print(13) is only going to show the column on the left after knitting.
```
We can also check what the first 5 rows look like:

```{r dataset-head}
# check head
datasaurus_dozen[1:5,]
```


### Exercise 2

First let's plot the data in the dino dataset:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

And next calculate the correlation between `x` and `y` in this dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```

### Exercise 3

<!-- Add code and narrative as needed. -->
<!-- Note that the R chunks are labelled with `plot-star` and `cor-star` to provide spaces to place the code for plotting and calculating the correlation coefficient. -->
<!-- To finish, clean up the narrative by removing these instructions. -->

Similarly, we first filter the observations belonging to the star dataset and store the filtering results into a new dataframe called `star_data`. Then, we use `ggplot` to visualize the data in points.

```{r plot-star, fig.height=6, fig.width=6}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x = x, y = y)) + 
  geom_point()
```

Again, we use the pipe operator `%>%` to send the dataframe `star_data` as the first argument to the `summarize` function and calculate the correlation coefficient. 

```{r cor-star}
star_data %>%
  summarize(r = cor(x, y))
```

Compared to the `r` of `dino`, the `r` of `star` is slightly bigger by 0.0015. 

### Exercise 4

<!-- Add code and narrative as needed. -->
<!-- Note that two R chunks are given but they are not labeled. -->
<!-- Use the convention from above to name them appropriately. -->

We filter the observations belonging to the `circle` dataset and store the filtering results into a new dataframe called `circle_data`. Then, we use `ggplot` to visualize the data in points.

```{r plot-circle, fig.height=6, fig.width=6}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

We can use the pipe operator `%>%` to send the dataframe `circle_data` as the first argument to the `summarize` function and calculate its correlation coefficient accordingly. 

```{r cor-circle}
circle_data %>%
  summarize(r = cor(x, y))
```

Compared to the `r` of `dino`, the `r` of `circle` is a little bit smaller by about 0.0038.

### Exercise 5

<!-- Add code and narrative as needed. -->
<!-- To add R chunks either type out the backticks, curly braces, and the letter `r` or use the Insert chunk button above, green C+. -->

First, let's use faceting to plot all datasets at once.

```{r plot-all, fig.height=12, fig.width=6}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset))+
  geom_point()+
  facet_wrap(~ dataset, ncol = 3) +
  theme(legend.position = "none")
```

Then, we use the `group_by` function to generate all the summary correlation coefficients.

```{r cor-all}
datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(r = cor(x, y)) #%>% print(13) I comment this out to show both rows after knitting
```

